Fast time-domain electromagnetic modelling in the frequency domain

### 2020-11-24 ManuscriptTime.tex
Make figwidth fit all formats

% Additional packages
% \usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage[draft, addedmarkup=bf]{changes}
% \usepackage[final]{changes}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

@@

% Change default colour of «changes» package
\definecolor{myred}{rgb}{.8,0,0}
\definecolor{mygreen}{rgb}{0,.5,0}
\setaddedmarkup{\color{myred}{#1}}
\setdeletedmarkup{\color{mygreen}{\itshape \sout{#1}}}

@@

There are three key-parameters that have to be defined: Minimum and maximum
frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
However, a simplified, layered model for the required survey setup and a fast
1D modeller makes it possible to estimate these parameters easily. To this end,
we created a graphical, interactive tool using \empymod, as shown in
Figure~\ref{fig:GUI-FFTLog-lin}.
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/GUI-FFTLog-lin}
  \caption{Interactive frequency selection for a user-provided layered model
    (the shown model parameters are described in the text). This example shows
    the impulse response at an offset of 5\,km, for which it uses FFTLog from
    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
  \label{fig:GUI-FFTLog-lin}
\end{figure}
%

@@

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\begin{figure*}
  \centering
  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-FFTLog-log}}%
  \label{fig:GUI-FFTLog-log}}\hfill
  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-DLF-lin}}%
  \label{fig:GUI-DLF-lin}}\\
  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
  scale. (b) Using DLF instead of FFTLog, but computing the same
  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
  response has comparable accuracy (compare the error to the error in Figure
  \ref{fig:GUI-FFTLog-lin}).}
  \label{fig:GUI-log-lin}
\end{figure*}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\begin{figure}
  \centering
  \includegraphics[width=.8\linewidth]{figures/multi-offset}
  \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
    $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
    \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
    responses, the black dots are the responses which are set to zero or
    interpolated. The black curves are the analytical responses.}
  \label{fig:multi-offset}
\end{figure}
%

@@

The first example is the inline electric field from a source at the origin
measured by an inline receiver with an offset of 900\,m in a homogeneous space
of $1\,\Omega\,$m. We chose this simple example to compare it with the
analytical solution and with previously published results. We used the
following values to define the required frequencies: $f_\mr{min}=0.05\,$Hz,
$f_\mr{max}=21\,$Hz, using FFTLog with 5 frequencies per decade. This results
in 14 frequencies to compute from 0.05\,Hz to 20.0\,Hz. The complete frequency
range for the transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz.
For the adaptive gridding the following inputs were used: $n_\delta=12$,
minimum cell width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
$\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
\num{46080} ($80\times24\times24$, for 20.0\,Hz) and \num{128000}
($80\times40\times40$, for 0.05\,Hz) cells. The run times for each frequency,
the corresponding number of cells, minimum cell width and computation domain
stretching factor are listed in Table~\ref{tbl:timefull}. The total run time to
compute this model was less than two minutes.
%
\begin{table}
\begin{minipage}{\columnwidth}
  \centering
  \caption{Run times per frequency for the homogeneous space example, with the
    corresponding number of cells and minimum cell width as well as the
    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
  \label{tbl:timefull}
  \begin{tabular}{S[table-format=2.4]rrrr}
    \toprule
    %
    {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
    {(Hz)}  & (s) &  & (m) & \\
    \midrule
    %
    20.0    &  3 & 80$\times$24$\times$24 & 20 & 1.26 \\
    12.6    &  6 & 96$\times$32$\times$32 & 20 & 1.17 \\
     7.98   &  8 & 96$\times$32$\times$32 & 20 & 1.20 \\
     5.03   &  7 & 96$\times$32$\times$32 & 20 & 1.23 \\
     3.18   &  7 & 80$\times$32$\times$32 & 24 & 1.25 \\
     2.00   &  5 & 80$\times$32$\times$32 & 30 & 1.21 \\
     1.26   &  4 & 64$\times$32$\times$32 & 37 & 1.21 \\
     0.798  &  5 & 64$\times$32$\times$32 & 40 & 1.23 \\
     0.503  &  5 & 64$\times$32$\times$32 & 40 & 1.26 \\
     0.318  &  5 & 64$\times$32$\times$32 & 40 & 1.28 \\
     0.200  & 10 & 64$\times$40$\times$40 & 40 & 1.27 \\
     0.126  & 10 & 64$\times$40$\times$40 & 40 & 1.30 \\
     0.0798 & 14 & 80$\times$40$\times$40 & 40 & 1.26 \\
     0.0503 & 15 & 80$\times$40$\times$40 & 40 & 1.28 \\
    %
    \bottomrule
  \end{tabular}
\end{minipage}
\end{table}
%

Figure \ref{fig:fullspace} (a) shows the frequency-domain result, where the
blue dots are the computed responses and the black dots correspond to the
interpolated values or the values set to zero.
%
\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/fullspace}
  \caption{(a) Frequency- and (b) time-domain results for the homogeneous space
    model. The red lines are the analytical solutions, the blue circles are the
    actually computed responses\deleted{ with \emg3d}, the black dots are the
    interpolated responses, and the dashed black line the obtained time-domain
    response. The errors are clipped for values smaller than 0.1\,\% and bigger
    than 100\,\%.}
  \label{fig:fullspace}
\end{figure*}
%
Most of the computed values stay below a relative error of 1\,\%, our chosen
adaptive gridding only starts to generate considerable errors at higher
frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
result, where the dashed black line is the result \replaced{obtained by
transformation of the frequency-domain response}{from \emg3d}, on
top of the red line which is the analytical result. The relative error is
mostly below 1\,\%, except for early times. However, for practical reasons that
is more than enough. Figure~\ref{fig:fullspace-log} shows the same on a
logarithmic scale, with times up to 10 seconds. It clearly shows that if later
times are required, we would need to adjust our Fourier transform parameters.
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/fullspace-log}
  \caption{Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic
    scale. To improve later times we would have to compute lower frequencies;
    to improve earlier times we would have to compute more frequencies per
    decade to get a better resolution.}
  \label{fig:fullspace-log}
\end{figure}
%
Note that for the gridding we chose $n_\delta=12$, which is very dense. This
was necessary because we are relatively close to the source. If the offsets of
interest are larger this factor can be lowered considerably; 3--4 is often
enough.

@@

The second example is a shallow marine, layered model with 200\,m of seawater
(3\,S/m) above a halfspace of 1\,S/m, an embedded target layer at 2\,km depth,
100\,m thick, with a conductivity of 0.02\,S/m. The source is located 20\,m
above the seafloor and the receivers are on the seafloor. We chose the
frequency range such that we can model offsets from 3 to 7 kilometers, with
$f_\mr{min}=0.007\,$Hz and $f_\mr{max}=32\,$Hz, using FFTLog with 5 frequencies
per decade. This results in computations for 19 frequencies from 0.008\,Hz to
31.8\,Hz. The complete frequency range for the transform includes 35
frequencies from $2\,10^{-5}$ to $126.4\,$Hz. For the adaptive gridding, we
used a cell width of 100\,m in the core domain and stretching outside up to a
factor 1.5, where the computation domain extends up to 50\,km in each
direction. This yielded grids between \num{204800} (higher frequencies) and
\num{245760} (lower frequencies) cells. The run times for each frequency and
their corresponding parameters are listed in Table~\ref{tbl:timemarine}.
%
\begin{table}
\begin{minipage}{\columnwidth}
  \centering
  \caption{Run times per frequency for the marine 1D example, with the
    corresponding number of cells and minimum cell width as well as the
    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
  \label{tbl:timemarine}
  \begin{tabular}{S[table-format=2.5]rrrr}
    \toprule
    %
    {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
    {(Hz)}  & (s) &  & (m) & \\
    \midrule
    %
    31.8     & 10 & 128$\times$40$\times$40 & 100 & 1.36 \\
    20.0     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
    12.6     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
     7.98    & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
     5.03    & 21 & 128$\times$40$\times$40 & 100 & 1.44 \\
     3.18    & 20 & 128$\times$40$\times$40 & 100 & 1.48 \\
     2.00    & 18 & 128$\times$40$\times$40 & 100 & 1.49 \\
     1.26    & 22 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.798   & 26 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.503   & 29 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.318   & 42 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.200   & 41 & 128$\times$40$\times$48 & 100 & 1.38 \\
     0.126   & 48 & 128$\times$40$\times$48 & 100 & 1.40 \\
     0.0798  & 54 & 128$\times$40$\times$48 & 100 & 1.41 \\
     0.0503  & 59 & 128$\times$40$\times$48 & 100 & 1.44 \\
     0.0318  & 74 & 128$\times$40$\times$48 & 100 & 1.44 \\
     0.0200  & 81 & 128$\times$40$\times$48 & 100 & 1.47 \\
     0.0126  & 96 & 128$\times$40$\times$48 & 100 & 1.48 \\
     0.00798 & 93 & 128$\times$40$\times$48 & 100 & 1.49 \\
    %
    \bottomrule
  \end{tabular}
\end{minipage}
\end{table}
%

Figure~\ref{fig:marine} shows the result for an offset of 5\,km, in (a) the
frequency and (b) the time domain. The recovered response with the 3D code
captures the airwave (first peak) and the subsurface (second peak) very
accurately. At later times the error starts to increase. We would need to
compute a few additional lower frequencies if we want to improve it.
%
\begin{figure*}
  \centering
  \includegraphics[width=.8\linewidth]{figures/marine}
  \caption{Electric inline response at an offset of 5\,km for a shallow marine,
    layered scenario. (a) Frequency-domain, where the blue circles are computed
    responses and the black dots are interpolated responses or responses set to
    zero. (b) Time-domain response.}
  \label{fig:marine}
\end{figure*}
%
In the frequency-domain plot, it can be seen that the high frequencies are not
computed very accurately, but without too much influence on the time-domain
response. These frequencies could be left out if an offset of 5\,km is the only
objective. However, we also want to retrieve shorter offsets from the same
computation, for which these frequencies are required.

Figure~\ref{fig:marine-multioffset} shows the time-domain responses of the same
model for offsets of 3, 5, and 7\,km, all obtained with the same
frequency-domain computations and the same frequencies for the Fourier
transform. The computation of these frequencies took less than 13 minutes and
it handles any offset between 3 and 7\,km. It can be seen that the chosen
frequency selection is sufficient for this offset range; again, more low
frequencies could be added to improve late-time values.
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/marine-multioffset}
  \caption{Time-domain responses for offsets of 3, 5, and 7\,km for the same
    model as shown in Figure~\ref{fig:marine}.}
  \label{fig:marine-multioffset}
\end{figure}
%

@@

Figure~\ref{fig:marine-wrong-x-y} shows this effect. It is the same model as in
the previous section; however, for the adaptive gridding in the horizontal
directions $\rho_\mr{ave}=1\,\Omega\,$m was used instead of
$\rho_\mr{ave}=\num{10000}\,\Omega\,$m. Having the boundaries too near in the
horizontal directions leads to worse results for most frequencies and entirely
wrong results for high frequencies. Comparison with the 1D result in the time
domain shows that it is the airwave whose amplitude is heavily overestimated.
%
\begin{figure*}
  \centering
  \includegraphics[width=.8\linewidth]{figures/marine-wrong-x-y}
  \caption{Same model as used for Figure~\ref{fig:marine}, but with the
    horizontal boundaries not far enough. Although the resulting time-domain
    result looks plausible, the comparison with the 1D result shows that it
    significantly overestimates the amplitude of the airwave.}
  \label{fig:marine-wrong-x-y}
\end{figure*}
%
It can be difficult to spot these errors in the time-domain result, as the
response looks plausible and only comparing it with the 1D result reveals that
it is actually wrong. A possibility to detect such problems for complicated
cases, where there is no semi-analytical result to compare with, is to compute
two or more models, moving the boundary. When the responses stop to change, one
can assume that the boundary is far enough. Another possibility is to look at
the amplitudes close to the boundaries and ensure that they are small enough.

@@

The final example consists of a resistive, three-dimensional block embedded in
the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
target has resistivity $\rho_\mr{tg}=100\,\Omega\,$m, the upper halfspace
corresponds to seawater with $\rho_\mr{sea}=0.3\,\Omega\,$m, and the lower
halfspace is the background with $\rho_\mr{bg}=1\,\Omega\,$m.
%
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/3d-model}
  \caption{Three-dimensional block embedded in the lower of two halfspaces. The
    100-m long, $x$-directed dipole source is located 50\,m above the seafloor
    at the origin, and the receiver is on the seafloor at an offset of 2\,km.}
  \label{fig:3d-model}
\end{figure}
%
The source is a 100-m long, $x$-directed dipole at the origin, 50\,m above the
seafloor, and we are using a step-off source function. The $x$-directed inline
receiver is at an offset of 2\,km. The dimension of the target cube is
$1.1\times1.0\times0.4$\,km, located 300\,m below the seafloor in the centre
between source and receiver.

>>>>

% Additional packages
% \usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{ifthen}
\usepackage[draft, addedmarkup=bf]{changes}
% \usepackage[final]{changes}
\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}

@@

% Figures
% To print widths put this in text:
% \the\textwidth
% \the\columnwidth
%
% Widths:   textwidth  columnwidth  tablewidth
%       1-c   504pt        252pt       504pt
%       2-c   504pt        240pt       240pt
%       ref   468pt        234pt       468pt
%
\ifthenelse{\boolean{GJ@referee}}{%
  \newcommand{\cwidth}{234pt}%
  \newcommand{\fwidth}{468pt}%
  \newcommand{\twidth}{468pt}}{%
  \newcommand{\fwidth}{504pt}%
  \ifthenelse{\boolean{@twocolumn}}{%
    \newcommand{\cwidth}{240pt}%
    \newcommand{\twidth}{240pt}}{%
    \newcommand{\cwidth}{252pt}%
    \newcommand{\twidth}{504pt}%
  }
}
\graphicspath{{./figures/}}

% Change default colour of «changes» package
\definecolor{myred}{rgb}{.8,0,0}
\definecolor{mygreen}{rgb}{0,.5,0}
% \setaddedmarkup{\color{myred}{#1}}
\setaddedmarkup{\color{myred}{#1}}
\setdeletedmarkup{\color{mygreen}{\sout{#1}}}

@@

There are three key-parameters that have to be defined: Minimum and maximum
frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
However, a simplified, layered model for the required survey setup and a fast
1D modeller makes it possible to estimate these parameters easily. To this end,
we created a graphical, interactive tool using \empymod, as shown in
Figure~\ref{fig:GUI-FFTLog-lin}.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{GUI-FFTLog-lin}
  \caption{Interactive frequency selection for a user-provided layered model
    (the shown model parameters are described in the text). This example shows
    the impulse response at an offset of 5\,km, for which it uses FFTLog from
    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
  \label{fig:GUI-FFTLog-lin}
\end{figure}
%

@@

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\begin{figure*}
  \centering
  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-FFTLog-log}}%
  \label{fig:GUI-FFTLog-log}}\hfill
  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-DLF-lin}}%
  \label{fig:GUI-DLF-lin}}\\
  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
  scale. (b) Using DLF instead of FFTLog, but computing the same
  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
  response has comparable accuracy (compare the error to the error in Figure
  \ref{fig:GUI-FFTLog-lin}).}
  \label{fig:GUI-log-lin}
\end{figure*}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{multi-offset}
  \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
    $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
    \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
    responses, the black dots are the responses which are set to zero or
    interpolated. The black curves are the analytical responses.}
  \label{fig:multi-offset}
\end{figure}
%

@@

The first example is the inline electric field from a source at the origin
measured by an inline receiver with an offset of 900\,m in a homogeneous space
of $1\,\Omega\,$m. We chose this simple example to compare it with the
analytical solution and with previously published results. We used the
following values to define the required frequencies: $f_\mr{min}=0.05\,$Hz,
$f_\mr{max}=21\,$Hz, using FFTLog with 5 frequencies per decade. This results
in 14 frequencies to compute from 0.05\,Hz to 20.0\,Hz. The complete frequency
range for the transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz.
For the adaptive gridding the following inputs were used: $n_\delta=12$,
minimum cell width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
$\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
\num{46080} ($80\times24\times24$, for 20.0\,Hz) and \num{128000}
($80\times40\times40$, for 0.05\,Hz) cells. The run times for each frequency,
the corresponding number of cells, minimum cell width and computation domain
stretching factor are listed in Table~\ref{tbl:timefull}. The total run time to
compute this model was less than two minutes.
%
\begin{table}
\begin{minipage}{\twidth}
  \centering
  \caption{Run times per frequency for the homogeneous space example, with the
    corresponding number of cells and minimum cell width as well as the
    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
  \label{tbl:timefull}
  \begin{tabular}{S[table-format=2.4]rrrr}
    \toprule
    %
    {Frequency} & Time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
    {(Hz)}  & (s) &  & (m) & \\
    \midrule
    %
    20.0    &  3 & 80$\times$24$\times$24 & 20 & 1.26 \\
    12.6    &  6 & 96$\times$32$\times$32 & 20 & 1.17 \\
     7.98   &  8 & 96$\times$32$\times$32 & 20 & 1.20 \\
     5.03   &  7 & 96$\times$32$\times$32 & 20 & 1.23 \\
     3.18   &  7 & 80$\times$32$\times$32 & 24 & 1.25 \\
     2.00   &  5 & 80$\times$32$\times$32 & 30 & 1.21 \\
     1.26   &  4 & 64$\times$32$\times$32 & 37 & 1.21 \\
     0.798  &  5 & 64$\times$32$\times$32 & 40 & 1.23 \\
     0.503  &  5 & 64$\times$32$\times$32 & 40 & 1.26 \\
     0.318  &  5 & 64$\times$32$\times$32 & 40 & 1.28 \\
     0.200  & 10 & 64$\times$40$\times$40 & 40 & 1.27 \\
     0.126  & 10 & 64$\times$40$\times$40 & 40 & 1.30 \\
     0.0798 & 14 & 80$\times$40$\times$40 & 40 & 1.26 \\
     0.0503 & 15 & 80$\times$40$\times$40 & 40 & 1.28 \\
    %
    \bottomrule
  \end{tabular}
\end{minipage}
\end{table}
%

Figure \ref{fig:fullspace} (a) shows the frequency-domain result, where the
blue dots are the computed responses and the black dots correspond to the
interpolated values or the values set to zero.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{fullspace}
  \caption{(a) Frequency- and (b) time-domain results for the homogeneous space
    model. The red lines are the analytical solutions, the blue circles are the
    actually computed responses\deleted{ with \emg3d}, the black dots are the
    interpolated responses, and the dashed black line the obtained time-domain
    response. The errors are clipped for values smaller than 0.1\,\% and bigger
    than 100\,\%.}
  \label{fig:fullspace}
\end{figure*}
%
Most of the computed values stay below a relative error of 1\,\%, our chosen
adaptive gridding only starts to generate considerable errors at higher
frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
result, where the dashed black line is the result \replaced{obtained by
transformation of the frequency-domain response}{from \emg3d}, on
top of the red line which is the analytical result. The relative error is
mostly below 1\,\%, except for early times. However, for practical reasons that
is more than enough. Figure~\ref{fig:fullspace-log} shows the same on a
logarithmic scale, with times up to 10 seconds. It clearly shows that if later
times are required, we would need to adjust our Fourier transform parameters.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{fullspace-log}
  \caption{Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic
    scale. To improve later times we would have to compute lower frequencies;
    to improve earlier times we would have to compute more frequencies per
    decade to get a better resolution.}
  \label{fig:fullspace-log}
\end{figure}
%
Note that for the gridding we chose $n_\delta=12$, which is very dense. This
was necessary because we are relatively close to the source. If the offsets of
interest are larger this factor can be lowered considerably; 3--4 is often
enough.

@@

The second example is a shallow marine, layered model with 200\,m of seawater
(3\,S/m) above a halfspace of 1\,S/m, an embedded target layer at 2\,km depth,
100\,m thick, with a conductivity of 0.02\,S/m. The source is located 20\,m
above the seafloor and the receivers are on the seafloor. We chose the
frequency range such that we can model offsets from 3 to 7 kilometers, with
$f_\mr{min}=0.007\,$Hz and $f_\mr{max}=32\,$Hz, using FFTLog with 5 frequencies
per decade. This results in computations for 19 frequencies from 0.008\,Hz to
31.8\,Hz. The complete frequency range for the transform includes 35
frequencies from $2\,10^{-5}$ to $126.4\,$Hz. For the adaptive gridding, we
used a cell width of 100\,m in the core domain and stretching outside up to a
factor 1.5, where the computation domain extends up to 50\,km in each
direction. This yielded grids between \num{204800} (higher frequencies) and
\num{245760} (lower frequencies) cells. The run times for each frequency and
their corresponding parameters are listed in Table~\ref{tbl:timemarine}.
%
\begin{table}
\begin{minipage}{\twidth}
  \centering
  \caption{Run times per frequency for the marine 1D example, with the
    corresponding number of cells and minimum cell width as well as the
    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
  \label{tbl:timemarine}
  \begin{tabular}{S[table-format=2.5]rrrr}
    \toprule
    %
    {Frequency} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
    {(Hz)}  & (s) &  & (m) & \\
    \midrule
    %
    31.8     & 10 & 128$\times$40$\times$40 & 100 & 1.36 \\
    20.0     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
    12.6     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
     7.98    & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
     5.03    & 21 & 128$\times$40$\times$40 & 100 & 1.44 \\
     3.18    & 20 & 128$\times$40$\times$40 & 100 & 1.48 \\
     2.00    & 18 & 128$\times$40$\times$40 & 100 & 1.49 \\
     1.26    & 22 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.798   & 26 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.503   & 29 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.318   & 42 & 128$\times$40$\times$48 & 100 & 1.36 \\
     0.200   & 41 & 128$\times$40$\times$48 & 100 & 1.38 \\
     0.126   & 48 & 128$\times$40$\times$48 & 100 & 1.40 \\
     0.0798  & 54 & 128$\times$40$\times$48 & 100 & 1.41 \\
     0.0503  & 59 & 128$\times$40$\times$48 & 100 & 1.44 \\
     0.0318  & 74 & 128$\times$40$\times$48 & 100 & 1.44 \\
     0.0200  & 81 & 128$\times$40$\times$48 & 100 & 1.47 \\
     0.0126  & 96 & 128$\times$40$\times$48 & 100 & 1.48 \\
     0.00798 & 93 & 128$\times$40$\times$48 & 100 & 1.49 \\
    %
    \bottomrule
  \end{tabular}
\end{minipage}
\end{table}
%

Figure~\ref{fig:marine} shows the result for an offset of 5\,km, in (a) the
frequency and (b) the time domain. The recovered response with the 3D code
captures the airwave (first peak) and the subsurface (second peak) very
accurately. At later times the error starts to increase. We would need to
compute a few additional lower frequencies if we want to improve it.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{marine}
  \caption{Electric inline response at an offset of 5\,km for a shallow marine,
    layered scenario. (a) Frequency-domain, where the blue circles are computed
    responses and the black dots are interpolated responses or responses set to
    zero. (b) Time-domain response.}
  \label{fig:marine}
\end{figure*}
%
In the frequency-domain plot, it can be seen that the high frequencies are not
computed very accurately, but without too much influence on the time-domain
response. These frequencies could be left out if an offset of 5\,km is the only
objective. However, we also want to retrieve shorter offsets from the same
computation, for which these frequencies are required.

Figure~\ref{fig:marine-multioffset} shows the time-domain responses of the same
model for offsets of 3, 5, and 7\,km, all obtained with the same
frequency-domain computations and the same frequencies for the Fourier
transform. The computation of these frequencies took less than 13 minutes and
it handles any offset between 3 and 7\,km. It can be seen that the chosen
frequency selection is sufficient for this offset range; again, more low
frequencies could be added to improve late-time values.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{marine-multioffset}
  \caption{Time-domain responses for offsets of 3, 5, and 7\,km for the same
    model as shown in Figure~\ref{fig:marine}.}
  \label{fig:marine-multioffset}
\end{figure}
%

@@

Figure~\ref{fig:marine-wrong-x-y} shows this effect. It is the same model as in
the previous section; however, for the adaptive gridding in the horizontal
directions $\rho_\mr{ave}=1\,\Omega\,$m was used instead of
$\rho_\mr{ave}=\num{10000}\,\Omega\,$m. Having the boundaries too near in the
horizontal directions leads to worse results for most frequencies and entirely
wrong results for high frequencies. Comparison with the 1D result in the time
domain shows that it is the airwave whose amplitude is heavily overestimated.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{marine-wrong-x-y}
  \caption{Same model as used for Figure~\ref{fig:marine}, but with the
    horizontal boundaries not far enough. Although the resulting time-domain
    result looks plausible, the comparison with the 1D result shows that it
    significantly overestimates the amplitude of the airwave.}
  \label{fig:marine-wrong-x-y}
\end{figure*}
%
It can be difficult to spot these errors in the time-domain result, as the
response looks plausible and only comparing it with the 1D result reveals that
it is actually wrong. A possibility to detect such problems for complicated
cases, where there is no semi-analytical result to compare with, is to compute
two or more models, moving the boundary. When the responses stop to change, one
can assume that the boundary is far enough. Another possibility is to look at
the amplitudes close to the boundaries and ensure that they are small enough.

@@

The final example consists of a resistive, three-dimensional block embedded in
the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
target has resistivity $\rho_\mr{tg}=100\,\Omega\,$m, the upper halfspace
corresponds to seawater with $\rho_\mr{sea}=0.3\,\Omega\,$m, and the lower
halfspace is the background with $\rho_\mr{bg}=1\,\Omega\,$m.
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{3d-model}
  \caption{Three-dimensional block embedded in the lower of two halfspaces. The
    100-m long, $x$-directed dipole source is located 50\,m above the seafloor
    at the origin, and the receiver is on the seafloor at an offset of 2\,km.}
  \label{fig:3d-model}
\end{figure}
%
The source is a 100-m long, $x$-directed dipole at the origin, 50\,m above the
seafloor, and we are using a step-off source function. The $x$-directed inline
receiver is at an offset of 2\,km. The dimension of the target cube is
$1.1\times1.0\times0.4$\,km, located 300\,m below the seafloor in the centre
between source and receiver.
